---
title: "TCGA CopyKit Analysis `r params$patient_code`"
author: "Anna K. Casasent, PhD"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
params:
  data_path: "/TCGA"
  disease_code: "LIHC"
  bam_folder: "marked"
  patient_code: "A3I0"
  sample_info_file: "TCGA-FV-A3I0-01A-13D-A91S-52_sample_info.tsv"
  cell_info_file: "TCGA-FV-A3I0-01A-13D-A91S-52_manifest.tsv"
  file_sep: "/"
  cor_res: 0.80
  k_res: 5
  output_folder: "00_Copykit_Basic_Analysis_Output"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = c("png", "pdf"), # which output figure type
                      dpi = 300,  # figure dot per inch 
                      echo = TRUE, # echos output of chunk in document
                      cache = FALSE, 
                      dep_auto = TRUE, # set it do try and figure out dependencies 
                      cache.comments = FALSE, # allows you change comments
                      tidy = TRUE, # Uses tidy to for a wrapper for code
                      tidy.opts = list(width.cutoff=60), # character number 
                      message = FALSE, # 
                      error = TRUE,    # ignores ERRORS still produces a PDF <- dangerous
                      include = TRUE,  # you can change this to be includes or excluded (to run or not to run)
                      eval = TRUE,     # says to evaluate each code chunk 
                      fig.height= 6,   # set standard figure height when included 
                      fig.width = 8,   # set standard figure width when included
                      fig.align = 'center' # centers the figure
                      )
```

# Analysis 
## Basic Processing 
### Load the library 
```{r library}
library(copykit)
library(params)
library(readxl)
library(tools)
```

### Analysis set up
These are the items that were set up in the params feature that need to be changed when we change tumors. 
```{r dir}
params$data_path
params$bam_folder
params$patient_code
params$sample_info_file
params$cell_info_file
params$file_sep
params$cor_res
params$k_res
```

### Read in files
Here we read in the cell information, the sample information, and our bam files. 
```{r files, message = FALSE}
input_dir <- paste(params$data_path,params$disease_code, params$patient_code,sep= params$file_sep)
output_dir <-  paste(params$data_path,params$disease_code, params$patient_code, params$output_folder, sep= params$file_sep)
  if(!file.exists(output_dir))
  { 
    print("created output directory")
    dir.create(output_dir)
  }
sample_info <- read.delim(paste(input_dir, params$sample_info_file, sep= params$file_sep))
cell_info <- read.delim(paste(input_dir,  params$cell_info_file, sep= params$file_sep))

# read in scDNA from 
tumor <- runVarbin(paste(input_dir,params$bam_folder,sep=params$file_sep),
                   remove_Y = TRUE, method="CBS", resolution="220kb")

passing_bams <- ncol(tumor)
```

Add details about the different samples. 
```{r add_TCGA_info}
# add the information in the cell_info file to the cells 
cell_info_order_sel <- cell_info[match(colnames(tumor), substring(cell_info$X.read1_filename, 0, (nchar(cell_info$X.read1_filename)-nchar(".fastq.gz")))),]
cell_info_order_sel$cell_name <- substring(cell_info_order_sel$X.read1_filename, 0, nchar(cell_info_order_sel$X.read1_filename)-nchar(".fastq.gz"))

#check that they match each other
all(cell_info_order_sel$cell_name == tumor$sample)

if(all(cell_info_order_sel$cell_name == tumor$sample))
{
  colData(tumor)$DNA.barcode <- cell_info_order_sel$DNA.barcode
  colData(tumor)$cell_id <- cell_info_order_sel$cell_id
  colData(tumor)$Sample <- cell_info_order_sel$Sample
  colData(tumor)$Row <- cell_info_order_sel$Row
  colData(tumor)$Col <- cell_info_order_sel$Col
  colData(tumor)$ploidy <- cell_info_order_sel$ploidy
  colData(tumor)$WELL <- cell_info_order_sel$WELL
  colData(tumor)$Barcode <- cell_info_order_sel$Barcode
  colData(tumor)$Cells1 <- cell_info_order_sel$Cells1
  colData(tumor)$Signal1 <- cell_info_order_sel$Signal1
  colData(tumor)$Confidence <- cell_info_order_sel$Confidence
  # add the information from the sample_info files
  colData(tumor)$sample_Short_ID <- sample_info$Short_ID
  colData(tumor)$sample_Single_Core_ID <- sample_info$Single_Core_ID
  colData(tumor)$sample_FAC_Tech <- sample_info$FAC_Tech
  colData(tumor)$sample_FACs_Date <- sample_info$FACs_Date
  colData(tumor)$sample_Dispense_Tech <- sample_info$Dispense_Tech
  colData(tumor)$sample_Dispense_Date <- sample_info$Dispense_Date
  colData(tumor)$sample_Library_Tech <- sample_info$Library_Tech
  colData(tumor)$sample_Date_plate <- sample_info$Date_plate
  colData(tumor)$sample_Chip.ID <- sample_info$Chip.ID
  colData(tumor)$sample_Times <- sample_info$Times
  colData(tumor)$sample_Tagmentation.time<- sample_info$Tagmentation.time
  colData(tumor)$sample_Cycles<- sample_info$Cycles
  colData(tumor)$sample_BP<- sample_info$BP
  colData(tumor)$sample_Barcode_Plate <- sample_info$Barcode_Plate
}

```

## QC Mertics
```{r Metric}
tumor <- runMetrics(tumor)
colData(tumor)$mean_bin <- apply(bincounts(tumor), 2, mean)
```

```{r Preposses_QC_Metrics_plots}
plotMetrics(tumor, metric = c("overdispersion", 
                              "breakpoint_count",
                              "reads_total",
                              "reads_assigned_bins"),
            label = "percentage_duplicates")
```

### Find Aneuploid
```{r mark_aneuploid}
# Mark aneuploid cells if they exist
tumor <- findAneuploidCells(tumor)
```

### Filter Low Quality Cells
The filtering for this samples used resolution `r params$cor_res` correlation with at least `r params$k_res`.
```{r mark_low_quality}
# Mark low-quality cells for filtering
tumor <- findOutliers(tumor, resolution = params$cor_res, k=params$k_res)
prefiltered_tumor <- tumor
saveRDS(prefiltered_tumor, file=paste(output_dir, paste(sample_info$DNA.barcode,"prefiltered_tumor.rds", sep="_"), sep=params$file_sep))
```

## Per-filtered Heatmap 
```{r Preposses_heatmap}
# Visualize cells labeled by filter and aneuploid status
plotHeatmap(tumor, label = c('outlier', 'is_aneuploid', 'ploidy', 'breakpoint_count', 'cell_corr_value','mean_bin'), row_split = 'outlier', n_threads=50)
```

# Filtering 

## Filter Outlier & Ploidy
Here we filter based on Computational outliers or if marked as diploid on a computational level. 

```{r Preposses_filtering}
# Remove cells marked as low-quality and/or aneuploid from the copykit object
num_outliers <- sum(colData(tumor)$outlier == TRUE)
tumor_rmoutliers <- tumor[,SummarizedExperiment::colData(tumor)$outlier == FALSE]
#saveRDS(tumor_rmoutliers, file=paste0())
saveRDS(tumor_rmoutliers, file=paste(output_dir, paste(sample_info$DNA.barcode,"tumor_rmoutliers_cor",params$cor_res,"k",params$k_res,".rds", sep="_"), sep=params$file_sep))

num_normals <- sum(colData(tumor)$is_aneuploid == FALSE)
tumor_rmnormals <- tumor[,SummarizedExperiment::colData(tumor)$is_aneuploid == TRUE]
saveRDS(tumor_rmoutliers, file=paste(output_dir, paste(sample_info$DNA.barcode,"tumor_rmnormals.rds", sep="_"),sep=params$file_sep))

tumor <- tumor[,SummarizedExperiment::colData(tumor)$outlier == FALSE]
tumor <- tumor[,SummarizedExperiment::colData(tumor)$is_aneuploid == TRUE]

saveRDS(tumor, file=paste(output_dir, paste(sample_info$DNA.barcode,"tumor_rmoutliers_cor",params$cor_res,"k",params$k_res,"rmnormals.rds", sep="_"), sep=params$file_sep))
```

# Summary 

This sample started with `r nrow(cell_info)` fastq files.

`r nrow(cell_info) - passing_bams` cells were dropped due to have too low coverage at the bam stage.

`r num_normals ` cells were were consider normal. 

`r num_outliers` cells were dropped due to having low correlation values with resolution `r params$cor_res` correlation with at least `r params$k_res`.

These might not add up because samples because the normal cells are calculated with considering quality.  

For this tumor we had `r ncol(tumor)` cell left after filtering both low quality and normal cells. 

The information for this samples rds files were saves using: `r sample_info$DNA.barcode` at each step. 
Please refer to this RMD/HTML report for questions. 

# Session Info

```{r sessioninfo}
sessionInfo()
```